---
title: "Bipartite - Why order of species matters?"
author: "Valentin Stefan"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(bipartite)
```

Below is the usual `Safariland` web that comes with the `bipartite` package. The species names come in a particular order. We will see that if we alter that order, we can get differences in some network metrics and this seems unexpected for me. **Why would the order of the species matter and then what is the correct order?**

```{r}
set.seed(1)
plotweb(Safariland)
```

```{r}
set.seed(1)
visweb(Safariland)
```

# Network level metrics

These are the network metrics that one can compute:

```{r}
set.seed(1)
net_1 <- networklevel(web = Safariland)
```

If we shuffle the insect species names in `Safariland`, we get differences in some network metrics. We do not alter the structure of the network, the abundance information stays unchanged, we just randomly input data in a different order (could be whatever order we typed in the species in the Excel file).

```{r}
# Shuffle columns to simulate whatever input order from an Excel file
set.seed(1)
Safariland_2 <- Safariland[, sample(colnames(Safariland))]
```

First, we plot the shuffled web to see that the order is changed but not abundance information:
```{r}
set.seed(1)
plotweb(Safariland_2)
```

```{r}
set.seed(1)
visweb(Safariland_2)
```

Compute again network metrics with the shuffled names:

```{r}
set.seed(1)
net_2 <- networklevel(web = Safariland_2)

# The two results are different!
identical(net_1, net_2)
```

Not all metrics change, but here are the ones where differences appear:
```{r}
net_df <- data.frame(net_1, net_2)
net_df$dif <- net_df$net_1 - net_df$net_2
net_df[net_df$dif != 0,]
```

Also, what if we change the order of the plant names and keep the order of the insects unaltered?

```{r}
# Shuffle columns
set.seed(1)
Safariland_3 <- Safariland[sample(rownames(Safariland)), ]

# Compute network metrics
set.seed(1)
net_3 <- networklevel(web = Safariland_3)

# The two results are different again!
identical(net_1, net_3)
```

Here are the differences:
```{r}
net_df <- data.frame(net_1, net_3)
net_df$dif <- net_df$net_1 - net_df$net_3
net_df[net_df$dif != 0,]
```

Order alphabetically the column names:

```{r}
# Sort
Safariland_4 <- Safariland[, sort(colnames(Safariland))]

# Compute network metrics
set.seed(1)
net_4 <- networklevel(web = Safariland_4)

# The two results are different again!
identical(net_1, net_4)
```

The differences:
```{r}
net_df <- data.frame(net_1, net_4)
net_df$dif <- net_df$net_1 - net_df$net_4
net_df[net_df$dif != 0,]
```

# Species level metrics

Same thing can happen also for species level metrics:

```{r}
set.seed(1)
sp_1 <- specieslevel(web = Safariland)
sp_2 <- specieslevel(web = Safariland_4)
identical(sp_1, sp_2)
all.equal(sp_1, sp_2)
```

```{r}
sessionInfo()
```

