---
title: "Bootstrap network metrics for comparison tests"
author: "by [Valentin Stefan](https://github.com/valentinitnelav) - last update `r format(Sys.time(), '%d %B %Y')`"
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r load-packages}
library(bipartite)
library(dplyr)
library(ggplot2)
```

# Read & prepare data
```{r}
dat <- read.csv("./estonia-case/data/interactions_with_plant_cover_2020_08_24.csv", sep = ";")
```

Is good practice to check for typos in species names
```{r, eval=FALSE}
dat$Plant_species %>% unique %>% sort
dat$Insect_species %>% unique %>% sort
```

Select only needed columns and expand rows
```{r}
df <- dat %>% select(Management_type, Plant_species, Insect_species, Num_Individuals)
# Repeat each row/interaction as many times as it was observed
ids <- rep(seq(1, nrow(df)), times = df$Num_Individuals)
df <- df[ids, ]

# Do the filtering here so that is not repeated in the for loop below
# unnecessarily N times.
lbl_1 <- "wooded meadow"
lbl_2 <- "alvar pasture"
df_1 <- df %>% filter(Management_type == lbl_1)
nrow(df_1)
df_2 <- df %>% filter(Management_type == lbl_2)
nrow(df_2)
```

# Visualize adjacency matrices

```{r}
mat_wm <- table(df_1$Plant_species, df_1$Insect_species)
visweb(mat_wm)
plotweb(mat_wm)
```

```{r}
mat_ap <- table(df_2$Plant_species, df_2$Insect_species)
visweb(mat_ap)
plotweb(mat_ap)
```

# Bootstrap network level metrics

Generate bootstrap distributions for the network metric.
```{r}
# Give network metric to compute (see bipartite::networklevel)
index <- "interaction evenness"

# Give a number of repetitions
N <- 100 # WARNING: number of repetitions influences p-value! Better choose 1000

# Prepare two empty vectors that will contain bootstrapped network metrics
net_1_boot_metrics <- net_2_boot_metrics <- rep(NA, N)

# For each repetition, sample with replacement from all interactions, separately
# for meadows and alvars, and compute a network metric.
for (i in 1:N){
  set.seed(i)
  net_1_boot_metrics[i] <- 
    df_1 %>% 
    slice_sample(prop = 1, replace = TRUE) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
  
  set.seed(i)
  net_2_boot_metrics[i] <- 
    df_2 %>% 
    slice_sample(prop = 1, replace = TRUE) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
}
```

Compute the observed metrics of the two networks.
```{r}
metric_net_1 <- df_1 %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
metric_net_1

metric_net_2 <- df_2 %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
metric_net_2
```

Plot the two distributions, their means and the observed metrics.
```{r}
observed_metrics <- data.frame(web = c(lbl_1, lbl_2),
                               obs = c(metric_net_1, metric_net_2))
observed_metrics
```

```{r}
boots_df <- data.frame(web = rep(c(lbl_1, lbl_2), each = N),
                       boots = c(net_1_boot_metrics, net_2_boot_metrics))

mean_boots <- boots_df %>% 
  group_by(web) %>% 
  summarise(mean = mean(boots),
            median = median(boots))
mean_boots
```

A problem here is that the observed metrics do not overlap with their corresponding the bootstrapped mean values.
```{r}
ggplot(data = boots_df,
       aes(x = boots,
           color = web)) +
  geom_density() +
  geom_vline(data = mean_boots,
             aes(xintercept = mean,
                 color = web),
             linetype = "dashed") +
  geom_vline(data = observed_metrics,
             aes(xintercept = obs,
                 color = web),
             linetype = "solid") +
  labs(x = index) +
  theme_bw()
```

