---
title: "Permutation test for two network metrics comparison"
author: "by [Valentin Stefan](https://github.com/valentinitnelav) - last update `r format(Sys.time(), '%d %B %Y')`"
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r load-packages}
library(bipartite)
library(dplyr)
```

# Read & prepare data
```{r}
dat <- read.csv("./estonia-case/data/interactions_with_plant_cover_2020_08_24.csv", sep = ";")
```

Is good practice to check for typos in species names
```{r, eval=FALSE}
dat$Plant_species %>% unique %>% sort
dat$Insect_species %>% unique %>% sort
```

Select only needed columns and expand rows
```{r}
df <- dat %>% select(Management_type, Plant_species, Insect_species, Num_Individuals)

# Repeat each row/interaction as many times as it was observed
ids <- rep(seq(1, nrow(df)), times = df$Num_Individuals)
df <- df[ids, ]

head(df)
```

# Permutations - all interactions from both networks
```{r}
# Give network metric to compute (see bipartite::networklevel)
index <- "H2"

# Give a number of repetitions
N <- 1000 # WARNING: number of repetitions influences p-value! Better choose 1000

lbl_1 <- "wooded meadow"
lbl_2 <- "alvar pasture"

# Prepare two empty vectors that will contain permuted network metrics
net_1_perm_metrics <- net_2_perm_metrics <- rep(NA, N)

for (i in 1:N){
  set.seed(i)
  
  # Create a new column with shuffled network labels
  df$lbl_shuffled <- sample(df$Management_type, replace = FALSE)
  
  # Using the shuffled network labels, re-construct a network 1 (wooded) and compute network metric
  net_1_perm_metrics[i] <- 
    df %>% 
    filter(lbl_shuffled == lbl_1) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
  
  # Using the shuffled network labels, re-construct a network 2 (alvar) and compute network metric
  net_2_perm_metrics[i] <- 
    df %>% 
    filter(lbl_shuffled == lbl_2) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
}
```

Compute the observed metrics of the two networks and their difference
```{r}
metric_net_1 <- df %>% 
    filter(Management_type == lbl_1) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
metric_net_1

metric_net_2 <- df %>% 
    filter(Management_type == lbl_2) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
metric_net_2

obs_dif_abs <- abs(metric_net_1 - metric_net_2)
obs_dif_abs
```

Plot the distribution of permuted differences and the observed difference (blue line, negative and positive). The red line corresponds to a difference of zero.
```{r}
dif <- (net_1_perm_metrics - net_2_perm_metrics)

hist(dif)
abline(v = 0, col = "red", lwd = 2)
abline(v = obs_dif_abs, col = "blue", lwd = 2)
abline(v = -obs_dif_abs, col = "blue", lwd = 2)
```

Get p-value
```{r}
# p-value
mean(abs(dif) >= obs_dif_abs)
```

# Permutations - sample bigger netowrk to the size of the samller one

```{r}
# Here we need to sample wooded meadow with the sample size given by the number
# of interactions from alvar pasture
n_inter_1 <- nrow(df %>% filter(Management_type == lbl_1)) # no. interactions for wooded meadow
n_inter_1
n_inter_2 <- nrow(df %>% filter(Management_type == lbl_2)) # no. interactions for alvar pasture
n_inter_2
spl_size <- min(n_inter_1, n_inter_2)

# Prepare empty vectors that will contain permuted network metrics
net_1_perm_metrics <- net_2_perm_metrics <- sample_metric <- rep(NA, N)

for (i in 1:N){
  set.seed(i)
  
  df_1 <- df %>% filter(Management_type == lbl_1) %>% slice_sample(n = spl_size) # wooded meadow
  df_2 <- df %>% filter(Management_type == lbl_2) # alvar pasture
  df_spl <- bind_rows(df_1, df_2)
  
  # Store the network metric of the sampled network
  sample_metric[i] <- df_1 %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
  
  # Create a new column with shuffled network labels
  df_spl$lbl_shuffled <- sample(df_spl$Management_type, replace = FALSE)
  
  # Using the shuffled network labels, re-construct a network 1 (wooded) and compute network metric
  net_1_perm_metrics[i] <- 
    df_spl %>% 
    filter(lbl_shuffled == lbl_1) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
  
  # Using the shuffled network labels, re-construct a network 2 (alvar) and compute network metric
  net_2_perm_metrics[i] <- 
    df_spl %>% 
    filter(lbl_shuffled == lbl_2) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index) 
}
```

Compute the observed metrics of the two networks and their difference. This time, for wooded meadow (the bigger network) we must estimate an observed metric as the average of all the sampled networks we took.
```{r}
# wooded meadow
metric_net_1 <- mean(sample_metric)
metric_net_1

# alvar pasture
metric_net_2 <- df %>% 
    filter(Management_type == lbl_2) %>% 
    select(Plant_species, Insect_species) %>% 
    table() %>% 
    bipartite::networklevel(index = index)
metric_net_2

obs_dif_abs <- abs(metric_net_1 - metric_net_2)
obs_dif_abs
```

Plot the distribution of permuted differences and the observed difference (blue line, negative and positive). The red line corresponds to a difference of zero.
```{r}
dif <- (net_1_perm_metrics - net_2_perm_metrics)

hist(dif)
abline(v = 0, col = "red", lwd = 2)
abline(v = obs_dif_abs, col = "blue", lwd = 2)
abline(v = -obs_dif_abs, col = "blue", lwd = 2)
```

Get p-value
```{r}
# p-value
mean(abs(dif) >= obs_dif_abs)
```